---
title: "CreateNinetyMinuteInductionSet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CreateNinetyMinuteInductionSet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
### load the rnaseq tools
```{r setup, include=FALSE}
library(brentlabRnaSeqTools)
```
### get the metadata biosample --> quality assessment
Note: right now, I haven't written in the ability to pass parameters to the server to filter the data before returning it. So, right now you need to pull all of the data and then filter it down if you only need a subset. There are functions to do this for the 90minuteInduction and environmental perturbation data included in the brentlabRnaSeqTools
```{r}
# api_url is a path that is loaded into the environment for you
combined_df = getMetadata(api_url)
```

```{r}
# a http 504 error means that the server has timed out. Please let me know when this happens b/c it will help with getting the data input/output settings right to prevent this error
raw_counts = getRawCounts(api_url)
```

```{r}
# note: grant_df is the 2016 grant summary specification of genotypes. it is automatically loaded with the brentlabRnaSeqTools package
induction_set = createNinetyMinuteInductionSet(combined_df, grant_df)
```

```{r}
passing_qc1_induction_set = qualityAssessmentFilter(induction_set)
```

```{r}
dds = deseqObjectWithProtocolSpecificSizeFactors(passing_qc1_induction_set, raw_counts )
```
### write to file to send to htcf for processing
```{r}
# make sure that scratch/* is in your .gitignore if you use this.
setwd("/home/chase/code/brentlab/brentlabRnaSeqTools/scratch")
write_rds(dds, "induction_set_for_DESeq_function.rds", compress="gz")
```



