---
title: "database_interaction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{database_interaction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup, include=TRUE}
library(brentlabRnaSeqTools)
library(RPostgres)
```


Below is an example of updating a record. Note that the slots of databse_info have been changed to non-functioning values. Check database_info itself for database options (you can do this by typing in database_info and tab or ?database_info). Please do not 'test' this code, or any of the code below, on a live database. However, you could create a playground sqlite database in memory and test these commands out there.

NOTE: there are syntax differences in the queries between sqlite and postgresql. The most significant is that postgresql is very picky about formating/capitalization and requires quotes around tables and maybe field names as a result. Check the postgresql documentation.
```{r update a record}
kn99_db = connectToDatabase(database_info$database_host, database_info$database_name, Sys.getenv('DB_USERNAME'), Sys.getenv('DB_PASSWORD'))

sql_stmt = glue_sql('UPDATE "bioSample"
                    SET genotype1 = \'CNAG_00830\'
                    WHERE "bioSample"."bioSampleNumber"
                      IN (2875, 2876)')
dbSendQuery(kn99_db, sql_stmt)

```

